---
title: "DSCI 5260 Group Project 6"
author: "Srikanth, Nishant, Soumya, Manimala"
date: "11/14/2021"
output:
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

### Loading Packages
```{r loadpackages, message=TRUE, warning=TRUE}

if(!require("pacman")) install.packages("pacman")
pacman::p_load(data.table,ggplot2,plyr,dplyr,corrplot,pROC,caret,glmnet,Rtsne,xgboost,PRROC,caTools,rpart,hrbrthemes,scales,wesanderson,smotefamily,rpart.plot,ROSE,MASS,naivebayes,FNN)
theme_set(theme_classic())
options(digits = 3)
set.seed(5738)
```

### Loading the dataset
```{r}
creditData <- read.csv('creditcard.csv')
```

### Exploring the dataset
```{r}
str(creditData)
colSums(is.na(creditData))
head(creditData)
summary(creditData)

table(creditData$Class)
prop.table(table(creditData$Class))
```

### Exploratory Data Analysis
### Distribution of Fraud Labels
```{r}

ggplot(data = creditData, 
       mapping = aes(x = factor(Class),
                     y= prop.table(stat(count)), 
                     fill = factor(Class))) + 
  geom_bar() + 
  theme(plot.title=element_text(size=25, face="bold", hjust = 0.5)) +
  theme_light() + 
  ggtitle("Distribution of Fraud Labels (%)") + 
  geom_bar(colour="black",position = "dodge") + 
  scale_x_discrete(labels = c("No Fraud", "Fraud"))+ 
  labs(x = 'Class', y = 'Percentage',fill="Class") + scale_y_continuous(labels = scales::percent) + 
  scale_fill_brewer(palette = "PRGn") 
```


### Distribution of Class vs Amount
```{r}
ggplot(data = creditData,
       aes(y=Amount,x=Class))+
  geom_point(col="slateblue3",size=2) + 
  facet_grid(~Class) + theme_light() + 
  labs(title="Distribution of Class vs Amount")
```


### Relationship between Class and Time
```{r}
creditData$Hours <- round(creditData$Time/3600)
## To view Time and Class relationship
# Chart 3
ggplot(data = creditData,
       aes(x=Time,fill=factor(Class)))+
  geom_histogram(col='black',bins=48) +  
  facet_grid(Class ~ ., scales = 'free_y') + 
  theme_light() + labs(title="Relationship between Class and Time",fill="Class") + scale_fill_brewer(palette = "RdBu")
```

### Relationship between Class and Hour (24 Hour)
```{r}
creditData$Hours[creditData$Hours > 23] <- creditData$Hours[creditData$Hours > 23] -24

ggplot(data = creditData,aes(x=Hours,fill=factor(Class)))+geom_histogram(col='black') +  facet_grid(Class ~ .,, scales = 'free_y') + theme_light() + labs(title="24 Hour Relationship between Class and Hour", fill="Class")

```




```{r}
# Move values of hours to time, to keep the structure simple and remove hours column
creditData$Time <- creditData$Hours
creditData <- subset(creditData,select=-Hours)
```

### Correlation plot of all the variables
```{r}
creditData$Class <- as.numeric(creditData$Class)
corr <- cor(creditData[],method="pearson")

library(RColorBrewer)
corrplot(corr, number.cex = .9, method = "circle", type = "full", tl.cex=0.8,tl.col = "black")
```



```{r}
#creditData <- creditData[,-1]

#Change 'Class' variable to factor
#creditData$Class <- as.factor(creditData$Class)
#levels(creditData$Class) <- c("Not Fraud", "Fraud")
```

### Data Preparation - Split the data
```{r}
set.seed(5738)
creditDataSplit <- createDataPartition(creditData$Class, p = .70, list = FALSE, times = 1)

creditData_train_Split <- creditData[ creditDataSplit,]
creditData_test_Split  <- creditData[-creditDataSplit,]
```


### Dataset after split
```{r}
table(creditData_train_Split$Class)
prop.table(table(creditData_train_Split$Class))

table(creditData_test_Split$Class)
prop.table(table(creditData_test_Split$Class))
```

### SMOTE application on Train & Test sets
```{r}
creditData_train_SM <- SMOTE(creditData_train_Split[,-30],creditData_train_Split$Class,K = 5)
creditData_train_SM <- creditData_train_SM$data # extract only the balanced dataset
creditData_train_SM$Class <- as.factor(creditData_train_SM$Class)
```


```{r}
table(creditData_train_SM$Class)
prop.table(table(creditData_train_SM$Class))
```

### Classification Model on the Train Dataset
```{r}
multi.class.model <- rpart(Class ~ ., data = creditData_train_Split)
rpart.plot(multi.class.model)
```

```{r}
rpart.rules(multi.class.model, cover = TRUE)

```

### CART on Train Dataset
```{r}
default.ct <- rpart(Class ~ ., data = creditData_train_Split, method = "class")
prp(default.ct, type = 1, extra = 1, under = TRUE, roundint = FALSE, 
    split.font = 2, varlen = -10, box.palette = "BuOr")
```

### Confusion Matrix on the Train Dataset
```{r}
default.ct.point.pred.train <- predict(default.ct, 
                                       data = creditData_train_Split, 
                                       type = "class")
confusionMatrix(default.ct.point.pred.train, as.factor(creditData_train_Split$Class))
```

### Predicting the values of Testdata using CART
```{r}
default.ct.point.pred.valid <- predict(default.ct, 
                                       newdata = creditData_test_Split, 
                                       type = "class")
confusionMatrix(default.ct.point.pred.valid, as.factor(creditData_test_Split$Class))
```

### ROC Curve - CART Algorithm
```{r}
roc_dtr <- roc.curve(creditData_test_Split$Class, default.ct.point.pred.valid,curve=TRUE)
roc_dtr
```
### Precision-Recall Curve - CART Algorithm
```{r}
pr_dtr <- pr.curve(creditData_test_Split$Class, default.ct.point.pred.valid,curve=TRUE)
pr_dtr
```

### Plot - Precision-Recall Curve
```{r}
plot(pr_dtr)
```

### Logistic Regression
```{r}
logit.reg <- glm(Class ~ ., data = creditData_train_Split, family = "binomial") 

options(scipen=999)
summary(logit.reg)
```

```{r}
exp(coef(logit.reg))
```


### Prediction on test Dataset - Logistic Regression
```{r}
logit.reg.pred <- predict(logit.reg, creditData_test_Split, type = "response")

t(t(head(logit.reg.pred, 10)))
```

### Confusion Matrix - Logistic Regression
```{r}
con_matrix_test <-table(creditData_test_Split$Class , logit.reg.pred > 0.5)
con_matrix_test
```

### ROC Curve - Logistic Regression
```{r}
roc_logit<- roc.curve(creditData_test_Split$Class, logit.reg.pred, curve = TRUE)
roc_logit
```
### Precision-Recall Curve - Logistic Regression
```{r}
pr_logit<- pr.curve(creditData_test_Split$Class, logit.reg.pred, curve = TRUE)
pr_logit
```

### Plot - Precision-Recall Curve
```{r}
plot(pr_logit)
```

### XGBoost Alogrithm
```{r}
labels <- creditData_train_Split$Class
y <- recode(labels, 'Not_Fraud' = 0, "Fraud" = 1)
```


```{r}
xgb <- xgboost(data = data.matrix(creditData_train_Split[,-31]), 
 label = y,
 eta = 0.1,
 gamma = 0.1,
 max_depth = 10, 
 nrounds = 300, 
 objective = "binary:logistic",
 colsample_bytree = 0.6,
 verbose = 0,
 nthread = 7,
 seed = 5738
)
```

### ROC Curve - XGBoost Algorithm
```{r}
xgb_pred <- predict(xgb, data.matrix(creditData_test_Split[,-31]))
roc.curve(creditData_test_Split$Class, xgb_pred, plotit = TRUE)
```

### Precision-Recall Curve - XGBoost Algorithm
```{r}
pr_xgb <-pr.curve(creditData_test_Split$Class, xgb_pred, curve = TRUE)
pr_xgb
```

Plot - Precision-Recall Curve
```{r}
plot(pr_xgb)
```


```{r}
creditData_train_Split$Class <- as.factor(creditData_train_Split$Class)
```

### Naive Bayes Classifier
```{r}
nb_model_1.1 <-naive_bayes(Class~.,data=creditData_train_Split)
summary(nb_model_1.1)
```

### Prediciton on test data using Naive Bayes Algorithm
```{r}
p_nb <- predict(nb_model_1.1, newdata = creditData_test_Split)
summary(p_nb)
```


### Confusion Matrix - Naive Bayes Classifier
```{r}
confusionMatrix <- confusionMatrix(table(p_nb ,creditData_test_Split$Class))
confusionMatrix
```

### ROC Curve - Naive Bayes Classifier
```{r}
roc.curve(creditData_test_Split$Class, p_nb, curve = TRUE)

```
### Precision-Recall Curve - Naive Bayes Classifier
```{r}
pr_nb <- pr.curve(creditData_test_Split$Class, p_nb, curve = TRUE)
pr_nb
```


### Plot - Precision-Recall Curve
```{r}
plot(pr_nb)
```

### KNN Algorithm
```{r}
nn <- knn(train = creditData_train_Split[,-31], test = creditData_test_Split[,-31], 
          cl = creditData_train_Split[,31], k = 5,prob = TRUE)
  # Nearest-neighbor Index (ratio of observed distance divided by the expected distance)
#attributes(nn)$prob
```

### ROC CUrve - KNN Algorithm
```{r}
roc(creditData_test_Split$Class, attributes(nn)$prob)
```

```{r}
plot(roc(creditData_test_Split$Class, attributes(nn)$prob),
     print.thres = T,
   print.auc=T)
```
### Precision-Recall Curve - KNN Algorithm
```{r}
pr_knn <- pr.curve(creditData_test_Split$Class, attributes(nn)$prob,curve = TRUE)
pr_knn
```


### Plot - Precision-Recall Curve
```{r}
plot(pr_knn)
```

